Galois Slicing as Differentiation

### 2025-07-11 main/introduction.tex
Drop indentation in  environment at end of 2.2.

Existing approaches to program slicing are often tied to particular programming languages or implementations.
In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
connection. Our main insight is that this approach to slicing can be seen as a generalised form of
differentiable programming, with approximations (slices) of elements playing the role of tangents.

>>>>

Existing approaches to program slicing are often tied to particular programming languages or implementations.
In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
backward slicing procedures generate Galois connections between these lattices. Our main insight is that this
approach can be seen as a generalised form of differentiable programming, with slices of elements playing the
role of tangents. Our categorical approach should provide a suitable setting for exploring data provenance for
a variety of programming languages, and is easily configured to use alternative approximation strategies,
including quantitative forms of slicing.
